Distributed Nonlinear Control Design using Separable Control Contraction Metrics

### 2017-06-23 R0V0.tex
Change title of the distribued optimisation section
Add Ian's changes on the paper

% Paragraph counter 
\newcounter{para}
\newcommand\mypara{\par \thesection.\refstepcounter{para}\thepara.\space}

@@

\mypara A forward complete solution $X^\ast(\cdot,x^\ast,u^\ast)$ to \eqref{eq:general system} is said to be  \emph{globally exponentially uniformly stabilizable} with \emph{rate} $\lambda>0$ if there exist a constant value $C>0$ and a feedback law $k^\ast:\mathbb{R}_{\geq0}\times\mathbb{R}^n\times\{X^\ast\}\times\{u^\ast\}\to\mathbb{R}^m$, denoted as $k^\ast$, such that the inequality
\begin{equation}\label{eq:global exponential uniform stabilizability}
	\left|X^\ast(t,x^\ast,u^\ast)-X(t,x,k^\ast)\right|\leq Ce^{-\lambda t}|x^\ast-x|
\end{equation}
holds, for every $t\geq0$, and for every initial condition $x\in\mathbb{R}^n$ (\cite{Manchester2014a}). Equivalently, if inequality~\eqref{eq:global exponential uniform stabilizability} holds, then the solution $X^\ast(\cdot,x^\ast,u^\ast)$ to \eqref{eq:general system} is said to be \emph{globally exponentially uniformly stable} for system \eqref{eq:general system} in closed loop with $k^\ast$

@@

\begin{figure}[htbp!]
	\centering
%	\includegraphics[]{}
	\resizebox{.4\columnwidth}{!}{\input{./imgs/Graphs.pdf_tex}}
	\caption{Illustration of the graphs $\mathscr{G}_p$ and $\mathscr{G}_c$. The solid lines represent the edges defining $\mathscr{G}_p$ while the dashed lines represent edges defining $\mathscr{G}_c$.}
	\label{fig:graph different illustration}
\end{figure}

@@

\mypara The importance of the stability of \eqref{eq:general system:differential} for system \eqref{eq:general system} can be understood as follows. Given a fixed control $u$ for system \eqref{eq:general system}, $\delta_u\equiv0$. If the Euclidean norm of every solution $\Delta_x(t,x,\delta_x,u,0)$ converges to zero exponentially, as the time tends to infinity, then every pair of solutions to \eqref{eq:general system} under the input $u$ converge to each other exponentially. The interested reader may address \cite{Lohmiller1998,Sontag2010} and references therein for further details.

@@

\begin{definition}[{\cite{Manchester2014a}}]\label{def:}
	A metric for system \eqref{eq:general system} is said to be a \emph{control contraction metric for system \eqref{eq:general system}} if there exists a constant value $\lambda>0$ such that the condition
	\begin{subequations}\label{eq:Arstein-Sontag}
		\begin{equation}\label{eq:}
			\delta_x\neq0,\quad\delta_x^\top M(x)B(x)=0
		\end{equation}
		implies that the inequality
		\begin{equation}\label{eq:Artstein-Sontag:inequality}
			\delta_x^\top(\dot{M}+A^\top M+MA)\delta_x<\ -2\lambda \delta_x^\top M\delta_x
		\end{equation}
		holds, where $\dot{M}:=\partial_{f+Bu}M$.
	\end{subequations}
\end{definition}
\mypara The set of equations \eqref{eq:Arstein-Sontag} is an adaptation of Artstein-Sontag's condition to the differential system~\eqref{eq:general system:differential}. Given a control contraction metric for system \eqref{eq:general system}, Finsler's lemma (cf. \cite[Lemma 11.1]{CalafioreGhaoui2014}) provides a stabilizing feedback law of the form $\delta_u=K(x,u)\delta_x$ for system \eqref{eq:general system:differential}, defined for every $(x,\delta_x,u)\in\mathbb{R}^n\times\mathbb{R}^n\times\mathbb{R}^m$, where $K:\mathbb{R}^n\times\mathbb{R}^m\to\mathbb{R}^{n\times m}$.

@@

\mypara Step 1 (Offline MI computation). Consider the change of variables $\eta=M\delta$ and define the matrix $W=M^{-1}$. The set of equations \eqref{eq:Arstein-Sontag} is equivalent (cf. \cite[Lemma 11.1]{CalafioreGhaoui2014}) to the existence of a bounded differentiable function $W:\mathbb{R}^n\to\mathbb{R}^{n\times n}$ such that $W(\cdot)=W(\cdot)^\top\succ0$ and a function $Y:\mathbb{R}^n\times\mathbb{R}^m\to\mathbb{R}^{m\times n}$ satisfying the following matrix inequality (MI)
\begin{equation}\label{eq:MI formulation}
	-\dot{W}+AW+WA^\top+BY+(BY)^\top+2\lambda W\prec0
\end{equation}
uniformly over $(x,u)\in\mathbb{R}^n\times\mathbb{R}^m$. Note that \eqref{eq:MI formulation} is linear on the matrix variables $W$ and $Y$, for every $(x,u)\in\mathbb{R}^n\times\mathbb{R}^m$. Consequently, the MI~\eqref{eq:MI formulation} can be solved with positive semidefinite programming techniques.

\mypara Once a solution to MI~\eqref{eq:MI formulation} has been computed, the function defined, for every $(x,\delta_x,u)\in\mathbb{R}^n\times\mathbb{R}^n\times\mathbb{R}^m$, by
	\begin{equation}\label{eq:differential feedback law}
		\delta_k=Y(x,u)W^{-1}(x)\delta_x:=K(x,u)\delta_x
	\end{equation}
	is a differential feedback law that renders the origin globally exponentially stable for system~\eqref{eq:general system:differential} in closed loop with $\delta_k$.

\mypara Step 2 (Online controller integration). The feedback law for system \eqref{eq:general system} can be obtained by integration as follows. Let $u^\ast:\mathbb{R}_{\geq0}\to\mathbb{R}^m$ be a control for system \eqref{eq:general system}, from Hopf-Rinow theorem (cf. \cite[Theorem 7.7]{Boothby1986}), for every pair of solutions $X(\cdot,x,u)$ and $X^\ast(\cdot,x^\ast,u^\ast)\in\mathbb{R}^n$, and at each time instant $t\geq0$,  there exist a smooth geodesic curve $\gamma:\{t\}\times [0,1]\to\mathbb{R}^n$ connecting them. This implies that the solution $k^\ast$ to the integral equation
\begin{equation}\label{eq:contracting feedback law}
	k^\ast(t,\sigma)=u^\ast(t)+\int_0^\sigma K(\gamma(t,s),k^\ast(t,s))\gamma_s(t,s)\,ds\;,
\end{equation}
where $s\in[0,1]$, is a feedback law rendering any forward complete solution $X^\ast$ to system~\eqref{eq:general system} globally exponentially stable.

@@

@@

Inspired by the notion of sum-separable Lyapunov functions (see \cite{Dirr2015}), we introduce the following class of control contraction metrics:

@@

\begin{theorem}\label{thm:main result}
	If there exists a solution to the matrix inequalities
	\begin{subequations}\label{eq:main result}
		\begin{equation}\label{eq:MI:SSCCM}
			-\dot{W}+AW+WA^\top+BY+(BY)^\top+2\lambda W\prec0
		\end{equation}
		with $W=\mathbin{\mathtt{diag}}(W_1,\ldots,W_N)$,
		\begin{equation}\label{eq:Ws:SSCCM}
			W_i\succ0, W_i\in\Pi_i,\forall i\in\mathbb{N}_{[1,N]}
		\end{equation}
		and
		\begin{equation}\label{eq:YinXi}
			Y\in\Xi.
		\end{equation}
	\end{subequations}
	Then, $W$ is a sum-separable control contraction metric for system~\eqref{eq:general system} and there exists a solution to Problem~\ref{problem formulation}. In particular, the graph $\mathscr{G}_c$ describing the communication network among different controllers specifies the set $\Xi$.
\end{theorem}

\begin{proof}[Proof of Theorem \ref{thm:main result}]
	By assumption, the functions $W:\mathbb{R}^n\to\mathbb{R}^{n\times n}$ and $Y:\mathbb{R}^n\times\mathbb{R}^m\to\mathbb{R}^{m\times n}$ satisfy the set of equations~\eqref{eq:main result}. Apply the coordinate change $\eta=M\delta$ and define the matrices $W=M^{-1}$ and $K=YW^{-1}$. Since $W$ is diagonal, the structure of $Y$ is preserved and, consequently, $K\in\Xi$.

@@

\section{Distributed Design of Distributed Controllers}\label{sec:distributed_design}

>>>>

% Paragraph counter 
\newcounter{para}
%\newcommand\mypara{\par \thesection.\refstepcounter{para}\thepara.\space}
\newcommand\mypara{\par}

@@

\mypara A forward complete solution $X^\ast(\cdot,x^\ast,u^\ast)$ to \eqref{eq:general system} is said to be  \emph{globally exponentially uniformly stabilizable} with \emph{rate} $\lambda>0$ if there exist a constant value $C>0$ and a feedback law $k^\ast:\mathbb{R}^n\times\{X^\ast\}\times\{u^\ast\}\to\mathbb{R}^m$, denoted as $k^\ast$, such that the inequality
\begin{equation}\label{eq:global exponential uniform stabilizability}
	\left|X^\ast(t,x^\ast,u^\ast)-X(t,x,k^\ast)\right|\leq Ce^{-\lambda t}|x^\ast-x|
\end{equation}
holds, for every $t\geq0$, and for every initial condition $x\in\mathbb{R}^n$ (\cite{Manchester2014a}). Equivalently, if inequality~\eqref{eq:global exponential uniform stabilizability} holds, then the solution $X^\ast(\cdot,x^\ast,u^\ast)$ to \eqref{eq:general system} is said to be \emph{globally exponentially uniformly stable} for system \eqref{eq:general system} in closed loop with $k^\ast$

@@

\begin{figure}
	\centering
%	\includegraphics[]{}
	%\resizebox{.4\columnwidth}{!}{\input{./imgs/Graphs.pdf_tex}}
	\caption{Illustration of the directed graphs representing the physical interaction between nodes $\mathscr{G}_p$, and the communication network $\mathscr{G}_c$. These may or may not be identical.}
	\label{fig:graph different illustration}
\end{figure}

@@

\mypara The importance of the stability of \eqref{eq:general system:differential} for system \eqref{eq:general system} can be understood as follows. Given a particular control signal $u(\cdot)$ for system \eqref{eq:general system}, $\delta_u\equiv0$. If the Euclidean norm of every solution $\Delta_x(t,x,\delta_x,u,0)$ converges to zero exponentially, as the time tends to infinity, then every pair of solutions to \eqref{eq:general system} under the input $u$ converge to each other exponentially. The interested reader may address \cite{Lohmiller1998,Sontag2010} and references therein for further details.

@@

\begin{definition}[{\cite{Manchester2014a}}]\label{def:}
	A metric for system \eqref{eq:general system} is said to be a \emph{control contraction metric for system \eqref{eq:general system}} if there exists a constant value $\lambda>0$ such that the condition
	\begin{subequations}\label{eq:Arstein-Sontag}
		\begin{equation}\label{eq:}
			\delta_x\neq0,\quad\delta_x^\top M(x)B(x)=0
		\end{equation}
		implies that the inequality
		\begin{equation}\label{eq:Artstein-Sontag:inequality}
			\delta_x^\top(\dot{M}+A^\top M+MA)\delta_x<\ -2\lambda \delta_x^\top M\delta_x
		\end{equation}
		holds, where $\dot{M}:=\partial_{f+Bu}M$.
	\end{subequations}
\end{definition}
\mypara The set of equations \eqref{eq:Arstein-Sontag} is an adaptation of Artstein-Sontag's condition to the differential system~\eqref{eq:general system:differential}. Given a control contraction metric for system \eqref{eq:general system}, Finsler's lemma (cf. \cite[Lemma 11.1]{CalafioreGhaoui2014}) provides a stabilizing feedback law of the form $\delta_u=K(x)\delta_x$ for system \eqref{eq:general system:differential}, defined for every $(x,\delta_x,u)\in\mathbb{R}^n\times\mathbb{R}^n\times\mathbb{R}^m$, where $K:\mathbb{R}^n\to\mathbb{R}^{n\times m}$.

@@

\mypara Step 1 (Offline MI computation). Consider the change of variables $\eta=M\delta$ and define the matrix $W=M^{-1}$. The set of equations \eqref{eq:Arstein-Sontag} is equivalent (cf. \cite[Lemma 11.1]{CalafioreGhaoui2014}) to the existence of a bounded differentiable function $W:\mathbb{R}^n\to\mathbb{R}^{n\times n}$ such that $W(\cdot)=W(\cdot)^\top\succ0$ and a function $Y:\mathbb{R}^n\to\mathbb{R}^{m\times n}$ satisfying the following matrix inequality (MI)
\begin{equation}\label{eq:MI formulation}
	-\dot{W}+AW+WA^\top+BY+(BY)^\top+2\lambda W\prec0
\end{equation}
uniformly over $(x,u)\in\mathbb{R}^n\times\mathbb{R}^m$. Note that \eqref{eq:MI formulation} is linear on the matrix variables $W$ and $Y$, for every $(x,u)\in\mathbb{R}^n\times\mathbb{R}^m$. Consequently, the MI~\eqref{eq:MI formulation} can be solved with positive semidefinite programming techniques.

\mypara Once a solution to MI~\eqref{eq:MI formulation} has been computed, the function defined, for every $(x,\delta_x,u)\in\mathbb{R}^n\times\mathbb{R}^n\times\mathbb{R}^m$, by
	\begin{equation}\label{eq:differential feedback law}
		\delta_k=Y(x)W^{-1}(x)\delta_x:=K(x)\delta_x
	\end{equation}
	is a differential feedback law that renders the origin globally exponentially stable for system~\eqref{eq:general system:differential} in closed loop with $\delta_k$.

\mypara Step 2 (Online controller integration). The feedback law for system \eqref{eq:general system} can be obtained by integration as follows. Let $u^\ast:\mathbb{R}_{\geq0}\to\mathbb{R}^m$ be a control for system \eqref{eq:general system}, from Hopf-Rinow theorem (cf. \cite[Theorem 7.7]{Boothby1986}), for every pair of solutions $X(\cdot,x,u)$ and $X^\ast(\cdot,x^\ast,u^\ast)\in\mathbb{R}^n$, and at each time instant $t\geq0$,  there exist a smooth geodesic curve $\gamma:\{t\}\times [0,1]\to\mathbb{R}^n$ connecting them. Then the differential feedback law $K(x)$ can be integrated along $\gamma$:
\begin{equation}\label{eq:contracting feedback law}
	k(x,=u^\ast(t)+\int_0^1 K(\gamma(t,s))\gamma_s(t,s)\,ds\;,
\end{equation}
where $s\in[0,1]$, is a feedback law rendering any forward complete solution $X^\ast$ to system~\eqref{eq:general system} globally exponentially stable.

Note about existence a.e.

Note about constant metric/constant Y case.

@@

This was used in \cite{wang_distributed_2017} to guarantee stability in distributed economic model predictive control.

@@

Inspired by the notion of sum-separable Lyapunov functions (see e.g. \cite{Dirr2015}), we introduce the following class of control contraction metrics:

@@

\begin{theorem}\label{thm:main result}
	If there exists a solution to the matrix inequalities
	\begin{subequations}\label{eq:main result}
		\begin{equation}\label{eq:MI:SSCCM}
			-\dot{W}+AW+WA^\top+BY+(BY)^\top+2\lambda W\prec0
		\end{equation}
		with $W=\mathbin{\mathtt{diag}}(W_1,\ldots,W_N)$,
		\begin{equation}\label{eq:Ws:SSCCM}
			W_i\succ0, W_i\in\Pi_i,\forall i\in\mathbb{N}_{[1,N]}
		\end{equation}
		and
		\begin{equation}\label{eq:YinXi}
			Y\in\Xi.
		\end{equation}
	\end{subequations}
	Then, $W$ is a separable control contraction metric for system~\eqref{eq:general system} and there exists a solution to Problem~\ref{problem formulation}. In particular, the graph $\mathscr{G}_c$ describing the communication network among different controllers specifies the set $\Xi$.
\end{theorem}

\begin{proof}[Proof of Theorem \ref{thm:main result}]
	By assumption, the functions $W:\mathbb{R}^n\to\mathbb{R}^{n\times n}$ and $Y:\mathbb{R}^n\times\mathbb{R}^m\to\mathbb{R}^{m\times n}$ satisfy the set of equations~\eqref{eq:main result}. Apply the coordinate change $\eta=M\delta$ and define the matrices $W=M^{-1}$ and $K=YW^{-1}$. Since $W$ is separable, the sparsity pattern and local dependence of $Y$ is preserved and, consequently, $K\in\Xi$.

@@

\section{Scalable Design of Distributed Controllers}\label{sec:distributed_design}
